\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
# Set your working directory

setwd("/home/guest/EDA_Spring2024/DamonLDanielsCourseProject")
getwd()

# Load your packages

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggridges)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(agricolae)
library(knitr)

# Set your ggplot theme

DLDtheme <- theme_classic(base_size = 8) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top") +
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

theme_set(DLDtheme)

```

```{r setup - Load Dataset, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
# Load your datasets

df = 
  read.csv(
  '/home/guest/EDA_Spring2024/DamonLDanielsCourseProject/Data/Raw_Data/NFWBS_PUF_2016_data.csv')

# The plyr library will be used to help us remap values
library(plyr)

# Codes on lines 49 through 1441 will change values and descriptors to match the CFPB NFWBS codebook. 

# This code maps values from the dummy variables to the descriptors from the codebook.
df$sample = revalue(factor(df$sample), c(
  `1` = "General population",
  `2` = "Age 62+ oversample",
  `3` = "Race/ethnicity and poverty oversample"
))
df$fpl = revalue(factor(df$fpl), c(
  `1` = "<100% FPL",
  `2` = "100%-199% FPL",
  `3` = "200%+ FPL"
))
df$SWB_1 = revalue(factor(df$SWB_1), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "1 Strongly disagree",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7 Strongly agree"
))
df$SWB_2 = revalue(factor(df$SWB_2), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "1 Strongly disagree",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7 Strongly agree"
))
df$SWB_3 = revalue(factor(df$SWB_3), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "1 Strongly disagree",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7 Strongly agree"
))
df$FWBscore = revalue(factor(df$FWBscore), c(
  `-4` = "Response not written to database",
  `-1` = "Refused"
))
df$FWB1_1 = revalue(factor(df$FWB1_1), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB1_2 = revalue(factor(df$FWB1_2), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB1_3 = revalue(factor(df$FWB1_3), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB1_4 = revalue(factor(df$FWB1_4), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB1_5 = revalue(factor(df$FWB1_5), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB1_6 = revalue(factor(df$FWB1_6), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FWB2_1 = revalue(factor(df$FWB2_1), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FWB2_2 = revalue(factor(df$FWB2_2), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FWB2_3 = revalue(factor(df$FWB2_3), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FWB2_4 = revalue(factor(df$FWB2_4), c(
  `-4` = "Response not written to database",
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FSscore = revalue(factor(df$FSscore), c(
  `-1` = "Refused"
))
df$FS1_1 = revalue(factor(df$FS1_1), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_2 = revalue(factor(df$FS1_2), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_3 = revalue(factor(df$FS1_3), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_4 = revalue(factor(df$FS1_4), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_5 = revalue(factor(df$FS1_5), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_6 = revalue(factor(df$FS1_6), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS1_7 = revalue(factor(df$FS1_7), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FS2_1 = revalue(factor(df$FS2_1), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FS2_2 = revalue(factor(df$FS2_2), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$FS2_3 = revalue(factor(df$FS2_3), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$SUBKNOWL1 = revalue(factor(df$SUBKNOWL1), c(
  `-1` = "Refused",
  `1` = "1 - Very low",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7 - Very high"
))
df$ACT1_1 = revalue(factor(df$ACT1_1), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$ACT1_2 = revalue(factor(df$ACT1_2), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Very little",
  `3` = "Somewhat",
  `4` = "Very well",
  `5` = "Completely"
))
df$FINGOALS = revalue(factor(df$FINGOALS), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$PROPPLAN_1 = revalue(factor(df$PROPPLAN_1), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$PROPPLAN_2 = revalue(factor(df$PROPPLAN_2), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$PROPPLAN_3 = revalue(factor(df$PROPPLAN_3), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$PROPPLAN_4 = revalue(factor(df$PROPPLAN_4), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$MANAGE1_1 = revalue(factor(df$MANAGE1_1), c(
  `-1` = "Refused",
  `1` = "Not applicable or never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$MANAGE1_2 = revalue(factor(df$MANAGE1_2), c(
  `-1` = "Refused",
  `1` = "Not applicable or never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$MANAGE1_3 = revalue(factor(df$MANAGE1_3), c(
  `-1` = "Refused",
  `1` = "Not applicable or never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$MANAGE1_4 = revalue(factor(df$MANAGE1_4), c(
  `-1` = "Refused",
  `1` = "Not applicable or never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$SAVEHABIT = revalue(factor(df$SAVEHABIT), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Disagree slightly",
  `4` = "Agree slightly",
  `5` = "Agree",
  `6` = "Strongly agree"
))
df$FRUGALITY = revalue(factor(df$FRUGALITY), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Disagree slightly",
  `4` = "Agree slightly",
  `5` = "Agree",
  `6` = "Strongly agree"
))
df$AUTOMATED_1 = revalue(factor(df$AUTOMATED_1), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes",
  `7` = "I do not have this type of account"
))
df$AUTOMATED_2 = revalue(factor(df$AUTOMATED_2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes",
  `7` = "I do not have this type of account"
))
df$ASK1_1 = revalue(factor(df$ASK1_1), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$ASK1_2 = revalue(factor(df$ASK1_2), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Seldom",
  `3` = "Sometimes",
  `4` = "Often",
  `5` = "Always"
))
df$SUBNUMERACY2 = revalue(factor(df$SUBNUMERACY2), c(
  `-1` = "Refused",
  `1` = "1 - Always prefer words",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6 - Always prefer numbers"
))
df$SUBNUMERACY1 = revalue(factor(df$SUBNUMERACY1), c(
  `-1` = "Refused",
  `1` = "1 - Not good at all",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6 - Extremely good"
))
df$CHANGEABLE = revalue(factor(df$CHANGEABLE), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Somewhat disagree",
  `4` = "Neither agree nor disagree",
  `5` = "Somewhat agree",
  `6` = "Agree",
  `7` = "Strongly agree"
))
df$GOALCONF = revalue(factor(df$GOALCONF), c(
  `-1` = "Refused",
  `1` = "Not at all confident",
  `2` = "Not very confident",
  `3` = "Somewhat confident",
  `4` = "Very confident"
))
df$FINKNOWL1 = revalue(factor(df$FINKNOWL1), c(
  `-1` = "Refused",
  `1` = "More than $102",
  `2` = "Exactly $102",
  `3` = "Less than $102"
))
df$FINKNOWL2 = revalue(factor(df$FINKNOWL2), c(
  `-1` = "Refused",
  `1` = "More than today",
  `2` = "Exactly the same",
  `3` = "Less than today"
))
df$FINKNOWL3 = revalue(factor(df$FINKNOWL3), c(
  `-1` = "Refused",
  `1` = "True",
  `2` = "False"
))
df$FK1correct = revalue(factor(df$FK1correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$FK2correct = revalue(factor(df$FK2correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$FK3correct = revalue(factor(df$FK3correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KHKNOWL1 = revalue(factor(df$KHKNOWL1), c(
  `-1` = "Refused",
  `1` = "Savings accounts",
  `2` = "Bonds",
  `3` = "Stocks"
))
df$KHKNOWL2 = revalue(factor(df$KHKNOWL2), c(
  `-1` = "Refused",
  `1` = "Savings accounts",
  `2` = "Bonds",
  `3` = "Stocks"
))
df$KHKNOWL3 = revalue(factor(df$KHKNOWL3), c(
  `-1` = "Refused",
  `1` = "Increase",
  `2` = "Decrease",
  `3` = "Stay the same"
))
df$KHKNOWL4 = revalue(factor(df$KHKNOWL4), c(
  `-1` = "Refused",
  `1` = "True",
  `2` = "False"
))
df$KHKNOWL5 = revalue(factor(df$KHKNOWL5), c(
  `-1` = "Refused",
  `1` = "True",
  `2` = "False"
))
df$KHKNOWL6 = revalue(factor(df$KHKNOWL6), c(
  `-1` = "Refused",
  `1` = "True",
  `2` = "False"
))
df$KHKNOWL7 = revalue(factor(df$KHKNOWL7), c(
  `-1` = "Refused",
  `1` = "Less than 5 years",
  `2` = "Between 5 and 10 years",
  `3` = "Between 10 and 15 years",
  `4` = "Never, you will continue to be in debt"
))
df$KHKNOWL8 = revalue(factor(df$KHKNOWL8), c(
  `-1` = "Refused",
  `1` = "They will rise",
  `2` = "They will fall",
  `3` = "They will stay the same",
  `4` = "There is no relationship between bond prices and the interest rate"
))
df$KHKNOWL9 = revalue(factor(df$KHKNOWL9), c(
  `-1` = "Refused",
  `1` = "True",
  `2` = "False"
))
df$KH1correct = revalue(factor(df$KH1correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH2correct = revalue(factor(df$KH2correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH3correct = revalue(factor(df$KH3correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH4correct = revalue(factor(df$KH4correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH5correct = revalue(factor(df$KH5correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH6correct = revalue(factor(df$KH6correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH7correct = revalue(factor(df$KH7correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH8correct = revalue(factor(df$KH8correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$KH9correct = revalue(factor(df$KH9correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$ENDSMEET = revalue(factor(df$ENDSMEET), c(
  `-1` = "Refused",
  `1` = "Not at all difficult",
  `2` = "Somewhat difficult",
  `3` = "Very difficult"
))
df$HOUSING = revalue(factor(df$HOUSING), c(
  `-1` = "Refused",
  `1` = "I own my home",
  `2` = "I rent",
  `3` = "I do not currently own or rent"
))
df$LIVINGARRANGEMENT = revalue(factor(df$LIVINGARRANGEMENT), c(
  `-1` = "Refused",
  `1` = "I am the only adult in the household",
  `2` = "I live with my spouse/partner/significant other",
  `3` = "I live in my parents' home",
  `4` = "I live with other family, friends, or roommates",
  `5` = "Some other arrangement"
))
df$HOUSERANGES = revalue(factor(df$HOUSERANGES), c(
  `-1` = "Refused",
  `1` = "Less than $300",
  `2` = "$300-499",
  `3` = "$500-749",
  `4` = "$750-999",
  `5` = "$1,000-1,499",
  `6` = "$1,500-1,999",
  `7` = "$2,000 or more",
  `98` = "I don't know",
  `99` = "Prefer not to say"
))
df$IMPUTATION_FLAG = revalue(factor(df$IMPUTATION_FLAG), c(
  `0` = "Not imputed",
  `1` = "Imputed"
))
df$VALUERANGES = revalue(factor(df$VALUERANGES), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `1` = "Less than $150,000",
  `2` = "$150,000-249,999",
  `3` = "$250,000-399,999",
  `4` = "$400,000 or more",
  `98` = "I don't know",
  `99` = "Prefer not to say"
))
df$MORTGAGE = revalue(factor(df$MORTGAGE), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `1` = "Less than $50,000",
  `2` = "$50,000-199,999",
  `3` = "$200,000 or more",
  `98` = "I don't know",
  `99` = "Prefer not to say"
))
df$SAVINGSRANGES = revalue(factor(df$SAVINGSRANGES), c(
  `-1` = "Refused",
  `1` = "0",
  `2` = "$1-99",
  `3` = "$100-999",
  `4` = "$1,000-4,999",
  `5` = "$5,000-19,999",
  `6` = "$20,000-74,999",
  `7` = "$75,000 or more",
  `98` = "I don't know",
  `99` = "Prefer not to say"
))
df$PRODHAVE_1 = revalue(factor(df$PRODHAVE_1), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_2 = revalue(factor(df$PRODHAVE_2), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_3 = revalue(factor(df$PRODHAVE_3), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_4 = revalue(factor(df$PRODHAVE_4), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_5 = revalue(factor(df$PRODHAVE_5), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_6 = revalue(factor(df$PRODHAVE_6), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_7 = revalue(factor(df$PRODHAVE_7), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_8 = revalue(factor(df$PRODHAVE_8), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODHAVE_9 = revalue(factor(df$PRODHAVE_9), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_1 = revalue(factor(df$PRODUSE_1), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_2 = revalue(factor(df$PRODUSE_2), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_3 = revalue(factor(df$PRODUSE_3), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_4 = revalue(factor(df$PRODUSE_4), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_5 = revalue(factor(df$PRODUSE_5), c(
  `0` = "No",
  `1` = "Yes"
))
df$PRODUSE_6 = revalue(factor(df$PRODUSE_6), c(
  `0` = "No",
  `1` = "Yes"
))
df$CONSPROTECT1 = revalue(factor(df$CONSPROTECT1), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Rarely",
  `3` = "Sometimes",
  `4` = "Often"
))
df$CONSPROTECT2 = revalue(factor(df$CONSPROTECT2), c(
  `-1` = "Refused",
  `1` = "Not at all familiar",
  `2` = "Somewhat familiar",
  `3` = "Very familiar"
))
df$CONSPROTECT3 = revalue(factor(df$CONSPROTECT3), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$EARNERS = revalue(factor(df$EARNERS), c(
  `-1` = "Refused",
  `1` = "One",
  `2` = "Two",
  `3` = "More than two"
))
df$VOLATILITY = revalue(factor(df$VOLATILITY), c(
  `-1` = "Refused",
  `1` = "Roughly the same each month",
  `2` = "Roughly the same most months, but some unusually high or low months during the year",
  `3` = "Often varies quite a bit from one month to the next"
))
df$SNAP = revalue(factor(df$SNAP), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes",
  `8` = "Not sure"
))
df$MATHARDSHIP_1 = revalue(factor(df$MATHARDSHIP_1), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$MATHARDSHIP_2 = revalue(factor(df$MATHARDSHIP_2), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$MATHARDSHIP_3 = revalue(factor(df$MATHARDSHIP_3), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$MATHARDSHIP_4 = revalue(factor(df$MATHARDSHIP_4), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$MATHARDSHIP_5 = revalue(factor(df$MATHARDSHIP_5), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$MATHARDSHIP_6 = revalue(factor(df$MATHARDSHIP_6), c(
  `-1` = "Refused",
  `1` = "Never",
  `2` = "Sometimes",
  `3` = "Often"
))
df$COLLECT = revalue(factor(df$COLLECT), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes",
  `8` = "Not sure"
))
df$REJECTED_1 = revalue(factor(df$REJECTED_1), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$REJECTED_2 = revalue(factor(df$REJECTED_2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$ABSORBSHOCK = revalue(factor(df$ABSORBSHOCK), c(
  `-1` = "Refused",
  `1` = "I am certain I could not come up with $2,000",
  `2` = "I could probably not come up with $2,000",
  `3` = "I could probably come up with $2,000",
  `4` = "I am certain I could come up with the full $2,000",
  `8` = "I don't know"
))
df$BENEFITS_1 = revalue(factor(df$BENEFITS_1), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$BENEFITS_2 = revalue(factor(df$BENEFITS_2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$BENEFITS_3 = revalue(factor(df$BENEFITS_3), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$BENEFITS_4 = revalue(factor(df$BENEFITS_4), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$BENEFITS_5 = revalue(factor(df$BENEFITS_5), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FRAUD2 = revalue(factor(df$FRAUD2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes",
  `8` = "Not sure"
))
df$COVERCOSTS = revalue(factor(df$COVERCOSTS), c(
  `-1` = "Refused",
  `1` = "Use savings or sell something you own",
  `2` = "Cut back or do without",
  `3` = "Earn more money",
  `4` = "Borrow money"
))
df$BORROW_1 = revalue(factor(df$BORROW_1), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$BORROW_2 = revalue(factor(df$BORROW_2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_1 = revalue(factor(df$SHOCKS_1), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_2 = revalue(factor(df$SHOCKS_2), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_3 = revalue(factor(df$SHOCKS_3), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_4 = revalue(factor(df$SHOCKS_4), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_5 = revalue(factor(df$SHOCKS_5), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_6 = revalue(factor(df$SHOCKS_6), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_7 = revalue(factor(df$SHOCKS_7), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_8 = revalue(factor(df$SHOCKS_8), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_9 = revalue(factor(df$SHOCKS_9), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_10 = revalue(factor(df$SHOCKS_10), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_11 = revalue(factor(df$SHOCKS_11), c(
  `0` = "No",
  `1` = "Yes"
))
df$SHOCKS_12 = revalue(factor(df$SHOCKS_12), c(
  `0` = "No",
  `1` = "Yes"
))
df$MANAGE2 = revalue(factor(df$MANAGE2), c(
  `-1` = "Refused",
  `1` = "Someone else takes care of all or most money matters in my household.",
  `2` = "Someone else and I take care of money matters in my household about the same.",
  `3` = "I take care of all or most money matters in my household."
))
df$PAIDHELP = revalue(factor(df$PAIDHELP), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$HSLOC = revalue(factor(df$HSLOC), c(
  `-1` = "Refused",
  `1` = "U.S. and territories",
  `2` = "Outside the U.S.",
  `8` = "I can't recall"
))
df$PAREDUC = revalue(factor(df$PAREDUC), c(
  `-1` = "Refused",
  `1` = "Less than high school",
  `2` = "High school degree/GED",
  `3` = "Some college/Associate",
  `4` = "Bachelor's degree",
  `5` = "Graduate/professional degree"
))
df$FINSOC2_1 = revalue(factor(df$FINSOC2_1), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_2 = revalue(factor(df$FINSOC2_2), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_3 = revalue(factor(df$FINSOC2_3), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_4 = revalue(factor(df$FINSOC2_4), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_5 = revalue(factor(df$FINSOC2_5), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_6 = revalue(factor(df$FINSOC2_6), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$FINSOC2_7 = revalue(factor(df$FINSOC2_7), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$OBJNUMERACY1 = revalue(factor(df$OBJNUMERACY1), c(
  `-1` = "Refused",
  `1` = "1%",
  `2` = "10%",
  `3` = "5%"
))
df$ON2correct = revalue(factor(df$ON2correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$ON1correct = revalue(factor(df$ON1correct), c(
  `0` = "No",
  `1` = "Yes"
))
df$MATERIALISM_1 = revalue(factor(df$MATERIALISM_1), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$MATERIALISM_2 = revalue(factor(df$MATERIALISM_2), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$MATERIALISM_3 = revalue(factor(df$MATERIALISM_3), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$CONNECT = revalue(factor(df$CONNECT), c(
  `-1` = "Refused"
))
df$HEALTH = revalue(factor(df$HEALTH), c(
  `-1` = "Refused",
  `1` = "Poor",
  `2` = "Fair",
  `3` = "Good",
  `4` = "Very good",
  `5` = "Excellent"
))
df$SCFHORIZON = revalue(factor(df$SCFHORIZON), c(
  `-1` = "Refused",
  `1` = "The next few months",
  `2` = "The next year",
  `3` = "The next few years",
  `4` = "The next 5 to 10 years",
  `5` = "Longer than 10 years"
))
df$DISCOUNT = revalue(factor(df$DISCOUNT), c(
  `-1` = "Refused",
  `1` = "$816 now",
  `2` = "$860 in three months"
))
df$MEMLOSS = revalue(factor(df$MEMLOSS), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$DISTRESS = revalue(factor(df$DISTRESS), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Agree",
  `5` = "Strongly agree"
))
df$SELFCONTROL_1 = revalue(factor(df$SELFCONTROL_1), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Not very well",
  `3` = "Very well",
  `4` = "Completely well"
))
df$SELFCONTROL_2 = revalue(factor(df$SELFCONTROL_2), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Not very well",
  `3` = "Very well",
  `4` = "Completely well"
))
df$SELFCONTROL_3 = revalue(factor(df$SELFCONTROL_3), c(
  `-1` = "Refused",
  `1` = "Not at all",
  `2` = "Not very well",
  `3` = "Very well",
  `4` = "Completely well"
))
df$OUTLOOK_1 = revalue(factor(df$OUTLOOK_1), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Somewhat disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Somewhat agree",
  `5` = "Strongly agree"
))
df$OUTLOOK_2 = revalue(factor(df$OUTLOOK_2), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Somewhat disagree",
  `3` = "Neither agree nor disagree",
  `4` = "Somewhat agree",
  `5` = "Strongly agree"
))
df$INTERCONNECTIONS_1 = revalue(factor(df$INTERCONNECTIONS_1), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_2 = revalue(factor(df$INTERCONNECTIONS_2), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_3 = revalue(factor(df$INTERCONNECTIONS_3), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_4 = revalue(factor(df$INTERCONNECTIONS_4), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_5 = revalue(factor(df$INTERCONNECTIONS_5), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_6 = revalue(factor(df$INTERCONNECTIONS_6), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_7 = revalue(factor(df$INTERCONNECTIONS_7), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_8 = revalue(factor(df$INTERCONNECTIONS_8), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_9 = revalue(factor(df$INTERCONNECTIONS_9), c(
  `0` = "No",
  `1` = "Yes"
))
df$INTERCONNECTIONS_10 = revalue(factor(df$INTERCONNECTIONS_10), c(
  `0` = "No",
  `1` = "Yes"
))
df$PEM = revalue(factor(df$PEM), c(
  `-1` = "Refused",
  `1` = "Strongly disagree",
  `2` = "Disagree",
  `3` = "Somewhat disagree",
  `4` = "Neither agree nor disagree",
  `5` = "Somewhat agree",
  `6` = "Agree",
  `7` = "Strongly agree"
))
df$HOUSESAT = revalue(factor(df$HOUSESAT), c(
  `-1` = "Refused",
  `1` = "Not at all satisfied",
  `2` = "Not very satisfied",
  `3` = "Somewhat satisfied",
  `4` = "Very satisfied"
))
df$SOCSEC1 = revalue(factor(df$SOCSEC1), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$SOCSEC2 = revalue(factor(df$SOCSEC2), c(
  `-3` = "Invalid Response",
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `62` = "62",
  `63` = "63",
  `64` = "64",
  `65` = "65",
  `66` = "66",
  `67` = "67",
  `68` = "68",
  `69` = "69",
  `70` = "70 and above"
))
df$SOCSEC3 = revalue(factor(df$SOCSEC3), c(
  `-3` = "Invalid Response",
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `61` = "61 and below",
  `62` = "62",
  `63` = "63",
  `64` = "64",
  `65` = "65",
  `66` = "66",
  `67` = "67",
  `68` = "68",
  `69` = "69",
  `70` = "70",
  `71` = "71 and above"
))
df$LIFEEXPECT = revalue(factor(df$LIFEEXPECT), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `0` = "0",
  `1` = "1",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7",
  `8` = "8",
  `9` = "9",
  `10` = "10",
  `11` = "11",
  `12` = "12",
  `15` = "15",
  `20` = "20",
  `22` = "22",
  `24` = "24",
  `25` = "25",
  `26` = "26",
  `28` = "28",
  `29` = "29",
  `30` = "30",
  `33` = "33",
  `35` = "35",
  `36` = "36",
  `37` = "37",
  `40` = "40",
  `41` = "41",
  `42` = "42",
  `43` = "43",
  `44` = "44",
  `45` = "45",
  `47` = "47",
  `50` = "50",
  `52` = "52",
  `54` = "54",
  `55` = "55",
  `56` = "56",
  `57` = "57",
  `59` = "59",
  `60` = "60",
  `62` = "62",
  `63` = "63",
  `64` = "64",
  `65` = "65",
  `66` = "66",
  `67` = "67",
  `68` = "68",
  `69` = "69",
  `70` = "70",
  `71` = "71",
  `72` = "72",
  `73` = "73",
  `75` = "75",
  `76` = "76",
  `77` = "77",
  `78` = "78",
  `79` = "79",
  `80` = "80",
  `81` = "81",
  `82` = "82",
  `83` = "83",
  `84` = "84",
  `85` = "85",
  `86` = "86",
  `87` = "87",
  `88` = "88",
  `89` = "89",
  `90` = "90",
  `91` = "91",
  `92` = "92",
  `93` = "93",
  `94` = "94",
  `95` = "95",
  `96` = "96",
  `97` = "97",
  `98` = "98",
  `99` = "99",
  `100` = "100"
))
df$HHEDUC = revalue(factor(df$HHEDUC), c(
  `-1` = "Refused",
  `1` = "Less than high school",
  `2` = "High school degree/GED",
  `3` = "Some college/Associate",
  `4` = "Bachelor's degree",
  `5` = "Graduate/professional degree"
))
df$KIDS_NoChildren = revalue(factor(df$KIDS_NoChildren), c(
  `-1` = "Refused",
  `0` = "Respondent financially supports children",
  `1` = "I have no children that I financially support"
))
df$KIDS_1 = revalue(factor(df$KIDS_1), c(
  `-1` = "Refused",
  `0` = "0",
  `1` = "1",
  `2` = "2+"
))
df$KIDS_2 = revalue(factor(df$KIDS_2), c(
  `-3` = "Invalid Response",
  `-1` = "Refused",
  `0` = "0",
  `1` = "1",
  `2` = "2+"
))
df$KIDS_3 = revalue(factor(df$KIDS_3), c(
  `-1` = "Refused",
  `0` = "0",
  `1` = "1",
  `2` = "2+"
))
df$KIDS_4 = revalue(factor(df$KIDS_4), c(
  `-1` = "Refused",
  `0` = "0",
  `1` = "1",
  `2` = "2+"
))
df$EMPLOY = revalue(factor(df$EMPLOY), c(
  `1` = "Self-employed",
  `2` = "Work full-time for an employer or the military",
  `3` = "Work part-time for an employer or the military",
  `4` = "Homemaker",
  `5` = "Full-time student",
  `6` = "Permanently sick, disabled or unable to work",
  `7` = "Unemployed or temporarily laid off",
  `8` = "Retired",
  `99` = "Refused"
))
df$EMPLOY1_1 = revalue(factor(df$EMPLOY1_1), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_2 = revalue(factor(df$EMPLOY1_2), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_3 = revalue(factor(df$EMPLOY1_3), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_4 = revalue(factor(df$EMPLOY1_4), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_5 = revalue(factor(df$EMPLOY1_5), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_6 = revalue(factor(df$EMPLOY1_6), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_7 = revalue(factor(df$EMPLOY1_7), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_8 = revalue(factor(df$EMPLOY1_8), c(
  `0` = "No",
  `1` = "Yes"
))
df$EMPLOY1_9 = revalue(factor(df$EMPLOY1_9), c(
  `0` = "No",
  `1` = "Yes"
))
df$RETIRE = revalue(factor(df$RETIRE), c(
  `-2` = "Question not asked because respondent not in item base",
  `-1` = "Refused",
  `1` = "Earlier than planned",
  `2` = "About when planned",
  `3` = "Later than planned"
))
df$MILITARY = revalue(factor(df$MILITARY), c(
  `-1` = "Refused",
  `0` = "No",
  `1` = "Yes"
))
df$Military_Status = revalue(factor(df$Military_Status), c(
  `-1` = "Refused",
  `1` = "Active military (active, reserve, natl guard)",
  `2` = "Veteran (veteran or retired)",
  `3` = "Spouse/dependent Active",
  `4` = "Spouse/dependent Veteran",
  `5` = "Neither Active nor Veteran"
))
df$agecat = revalue(factor(df$agecat), c(
  `1` = "18-24",
  `2` = "25-34",
  `3` = "35-44",
  `4` = "45-54",
  `5` = "55-61",
  `6` = "62-69",
  `7` = "70-74",
  `8` = "75+"
))
df$generation = revalue(factor(df$generation), c(
  `1` = "Pre-Boomer",
  `2` = "Boomer",
  `3` = "Gen X",
  `4` = "Millennial"
))
df$PPEDUC = revalue(factor(df$PPEDUC), c(
  `1` = "Less than high school",
  `2` = "High school degree/GED",
  `3` = "Some college/Associate",
  `4` = "Bachelor's degree",
  `5` = "Graduate/professional degree"
))
df$PPETHM = revalue(factor(df$PPETHM), c(
  `1` = "White, Non-Hispanic",
  `2` = "Black, Non-Hispanic",
  `3` = "Other, Non-Hispanic",
  `4` = "Hispanic"
))
df$PPGENDER = revalue(factor(df$PPGENDER), c(
  `1` = "Male",
  `2` = "Female"
))
df$PPHHSIZE = revalue(factor(df$PPHHSIZE), c(
  `1` = "1",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5+"
))
df$PPINCIMP = revalue(factor(df$PPINCIMP), c(
  `1` = "Less than $20,000",
  `2` = "$20,000 to $29,999",
  `3` = "$30,000 to $39,999",
  `4` = "$40,000 to $49,999",
  `5` = "$50,000 to $59,999",
  `6` = "$60,000 to $74,999",
  `7` = "$75,000 to $99,999",
  `8` = "$100,000 to $149,999",
  `9` = "$150,000 or more"
))
df$PPMARIT = revalue(factor(df$PPMARIT), c(
  `1` = "Married",
  `2` = "Widowed",
  `3` = "Divorced/Separated",
  `4` = "Never married",
  `5` = "Living with partner"
))
df$PPMSACAT = revalue(factor(df$PPMSACAT), c(
  `0` = "Non-Metro",
  `1` = "Metro"
))
df$PPREG4 = revalue(factor(df$PPREG4), c(
  `1` = "Northeast",
  `2` = "Midwest",
  `3` = "South",
  `4` = "West"
))
df$PPREG9 = revalue(factor(df$PPREG9), c(
  `1` = "New England",
  `2` = "Mid-Atlantic",
  `3` = "East-North Central",
  `4` = "West-North Central",
  `5` = "South Atlantic",
  `6` = "East-South Central",
  `7` = "West-South Central",
  `8` = "Mountain",
  `9` = "Pacific"
))
df$PPT01 = revalue(factor(df$PPT01), c(
  `0` = "0",
  `1` = "1+"
))
df$PPT25 = revalue(factor(df$PPT25), c(
  `0` = "0",
  `1` = "1+"
))
df$PPT612 = revalue(factor(df$PPT612), c(
  `0` = "0",
  `1` = "1+"
))
df$PPT1317 = revalue(factor(df$PPT1317), c(
  `0` = "0",
  `1` = "1+"
))
df$PPT18OV = revalue(factor(df$PPT18OV), c(
  `1` = "1",
  `2` = "2",
  `3` = "3",
  `4` = "4+"
))
df$PCTLT200FPL = revalue(factor(df$PCTLT200FPL), c(
  `-5` = "County is not known",
  `0` = "Less than 40% of county population below 200% of poverty level",
  `1` = "40% or more of county population below 200% of poverty level"
))

write.csv(df, row.names = FALSE,
          file = "./Data/Processed_Data/NFWBS_PUF_2016_dataPROCESSED.csv")

```

```{r Data Wrangling, include=FALSE}

# Goal: wrangle data containing information about (1) income levels, (2) self-reported levels of financial literacy, (3) self-reported levels of financial knowledge, and (4) Two separate scales that gauge correctness of self-reported levels of financial knowledge

# Relevant Columns for Wrangling:
  # Column  1     (PUF ID - Public Use File ID)
  # Column  2     (sample)
  # Column  3     (fpl - Federal Poverty Status)
  # Columns 4-6   (SWB - Subjective Well-Being Scores, 3 separate questions)
  # Column  7     (FWBscore - Financial Well-Being Scale Score, Total)
  # Column  18    (FSscore - Financial Skill Scale Scores)
  # Column  51    (LMscore - Lusardi & Mitchell Financial Knowledge Scale Score)
  # Column  58    (KHscore - Knoll & Houts Financial Knowledge Scale Score) 

# Create New Dataframe

dim(df)

df2 <- select(df, PUF_ID:SWB_3, FWBscore, FSscore, LMscore, KHscore)
dim(df2)

# (1) Remove extra text from all cells that also contain number values:
# (2) Replace all cells containing [a] "Response not written to database," and [b] "Refused" with "0"


df3 <- df2 
df3 <- df3 %>%
  mutate_all(as.character)
df3[df3 == "7 Strongly agree"] <- 7
df3[df3 == "1 Strongly disagree"] <- 1
df3[df3 == "Response not written to database"] <- 0
df3[df3 == "Refused"] <- 0
df3 <- df3 %>% mutate_if(is.character,as.factor)

# Save processed data 

write.csv(df2, row.names = FALSE,
          file = "./Data/Processed_Data/NFWBS_PUF_2016_data_select_PROCESSED.csv")

```

# Rationale and Research Questions

# The objective of the Consumer Financial Protection Bureau (CFPB) is to enhance financial well-being among Americans. Financial literacy education plays a critical role in this effort. Unfortunately, in attempting to evaluate the effectiveness of such efforts, it became clear that the consumer finance and consumer capability fields lacked any accepted standard definition or measure of financial well-being initiatives (CFPB 2017). CFPB decided to engage in year-long research to develop an evidence-based and consumer-driven definition of financial well-being, create a validated scale from this data -- the CFPB Financial Well-Being Scale -- and then conduct a national survey using the scale, the CFPB National Financial Well-Being Survey.

# The Research Question is as follows:

# What associations can be drawn between an (1) an individual’s or household’s federal poverty level (FPL), (2) self-reported assessments of financial well-being, and (3) measures of financial knowledge?


\newpage

# Dataset Information

# Conducted in English and Spanish between October 27, 2016 and December 5, 2016, 6,394 surveys were completed in total: 5,395 from the general population sample and 999 from an oversample of adults aged 62 and older. These samples are representative of the U.S. adult population plus the District of Columbia. This survey was fielded using the GfK Knowledge Panel, which recruited individuals using address-based sampling and dual-frame landline and cell phone random digit dialing methods.

# A total of 217 variables were among the complete original dataset. For this specfic project, these have been reduced to 54 total variables that represent the most pertinent partcipant responses that will provide answers to the research question.


\newpage

# Exploratory Analysis 

# View dimensions, column names, and variable type for each column:

```{r Summary Data, echo=FALSE}

class(df2)
colnames(df2)
str(df2)
dim(df2)
length(df2)
summary(df2)

```


\newpage

#  Data Visualizations, by Federal Poverty Level (FPL)

```{Data Visualizations, by Federal Poverty Level, echo=FALSE}

summary(df2$fpl)

# Subjective Well-Being (SWB) measures by Federal Poverty Level (FPL):

mean_SWB_1 <- mean(as.integer(df3$SWB_1))
print(mean_SWB_1)
mean_SWB_2 <- mean(as.integer(df3$SWB_2))
print(mean_SWB_2)
mean_SWB_3 <- mean(as.integer(df3$SWB_3))
print(mean_SWB_3)

df3_mean_SWB.processed <- 
  df3 %>%
  group_by(SWB_1, SWB_2, SWB_3)
df3_mean_SWB.processed <- 
  df3 %>%
  summarise(
    mean_SWB_1 = mean(as.integer(df3_mean_SWB.processed$SWB_1)),
    mean_SWB_2 = mean(as.integer(df3_mean_SWB.processed$SWB_2)),
    mean_SWB_3 = mean(as.integer(df3_mean_SWB.processed$SWB_3)))

df3_fpl1.processed <- 
  df3 %>%
  filter(fpl == "200%+ FPL") %>%
  group_by(SWB_1, SWB_2, SWB_3)
  df3_fpl1.processed <- 
  df3 %>%
  summarise(
    mean_SWB_1 = mean(as.integer(df3_fpl1.processed$SWB_1)),
    mean_SWB_2 = mean(as.integer(df3_fpl1.processed$SWB_2)),
    mean_SWB_3 = mean(as.integer(df3_fpl1.processed$SWB_3)))

df3_fpl2.processed <- 
  df3 %>%
  filter(fpl == "100%-199% FPL") %>%
  group_by(SWB_1, SWB_2, SWB_3)
  df3_fpl2.processed <- 
  df3 %>%
  summarise(
    mean_SWB_1 = mean(as.integer(df3_fpl2.processed$SWB_1)),
    mean_SWB_2 = mean(as.integer(df3_fpl2.processed$SWB_2)),
    mean_SWB_3 = mean(as.integer(df3_fpl2.processed$SWB_3)))

df3_fpl3.processed <- 
  df3 %>%
  filter(fpl == "<100% FPL") %>%
  group_by(SWB_1, SWB_2, SWB_3)
df3_fpl3.processed <- 
  df3 %>%
  summarise(
    mean_SWB_1 = mean(as.integer(df3_fpl3.processed$SWB_1)),
    mean_SWB_2 = mean(as.integer(df3_fpl3.processed$SWB_2)),
    mean_SWB_3 = mean(as.integer(df3_fpl3.processed$SWB_3)))

# Merge Processed dataframes into list:

mean_SWB_by_fpl <-
  list(
    df3_fpl3.processed, df3_fpl2.processed, df3_fpl1.processed, df3_mean_SWB.processed)

Reduce(function(x, y) merge(x, y, all=TRUE), mean_SWB_by_fpl)

mean_SWB_by_fpl2 <- data.frame(
  matrix(
    unlist(
      mean_SWB_by_fpl), nrow=length(mean_SWB_by_fpl), byrow=TRUE)) 

colnames(mean_SWB_by_fpl2) <- c("SWB_1", "SWB_2", "SWB_3")

mean_SWB_by_fpl2$meanSWB <- c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")

mean_SWB_by_fpl2 <- select(mean_SWB_by_fpl2, meanSWB,SWB_1,SWB_2,SWB_3)

# Plots

fplLevels <- ggplot(mean_SWB_by_fpl2) +
  geom_col(aes(
    x=factor(
      meanSWB, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=SWB_1, group = 1, color = meanSWB, fill = meanSWB))
print(fplLevels)

fplLevels <- fplLevels +
  theme(legend.position = "none") +
  labs(title = "", x = "", y = "I am satisfied with my life")
print(fplLevels)

fplLevels2 <- ggplot(mean_SWB_by_fpl2) +
  theme(legend.position = "none") +
  geom_col(aes(
    x=factor(
      meanSWB, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=SWB_2, group = 1, color = meanSWB, fill = meanSWB))
print(fplLevels2)

fplLevels2 <- fplLevels2 +
  theme(legend.position = "none") +
  labs(title = "", x = "", y = "I am optimistic about my future")
print(fplLevels2)

fplLevels3 <- ggplot(mean_SWB_by_fpl2) +
  theme(legend.position = "bottom") +
  geom_col(aes(
    x=factor(
      meanSWB, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=SWB_3, group = 1, color = meanSWB, fill = meanSWB))
print(fplLevels3)

fplLevels3 <- fplLevels3 +
  labs(title = "", x = "", y = "If I work hard today, I will be more successful in the future")
print(fplLevels3)

SWB_by_FPL <- plot_grid(fplLevels, fplLevels2, fplLevels3,
  nrow = 1, align = "h", rel_heights = c(1, 1))
print(SWB_by_FPL)

# Mean FWB Score measures by Federal Poverty Level

mean_FWBscore <- mean(as.integer(df3$FWBscore))
print(mean_FWBscore)

df3_meanFWBscore.processed <- 
  df3 %>%
  select(FWBscore)
df3_meanFWBscore.processed <-
  df3 %>%
  summarise(
    mean_FWBscore = mean(as.integer(df3$FWBscore)))

meanFWBscore_byFPL1 <- 
  df3 %>%
  filter(fpl == "200%+ FPL") %>%
  select(FWBscore)
meanFWBscore_byFPL1 <- 
  meanFWBscore_byFPL1 %>%
  summarise(
    mean_FWBscore = mean(as.integer(meanFWBscore_byFPL1$FWBscore)))

meanFWBscore_byFPL2 <- 
  df3 %>%
  filter(fpl == "100%-199% FPL") %>%
  select(FWBscore)
meanFWBscore_byFPL2 <- 
  meanFWBscore_byFPL2 %>%
  summarise(
    mean_FWBscore = mean(as.integer(meanFWBscore_byFPL2$FWBscore)))

meanFWBscore_byFPL3 <- 
  df3 %>%
  filter(fpl == "<100% FPL") %>%
  select(FWBscore)
meanFWBscore_byFPL3 <- 
  meanFWBscore_byFPL3 %>%
  summarise(
    mean_FWBscore = mean(as.integer(meanFWBscore_byFPL3$FWBscore)))

mean_FWB_by_fpl.list <-
  list(
    meanFWBscore_byFPL3, meanFWBscore_byFPL2, meanFWBscore_byFPL1, df3_meanFWBscore.processed)

Reduce(function(x, y) merge(x, y, all=TRUE), mean_FWB_by_fpl.list)

mean_FWB_by_fpl.list2 <- data.frame(
  matrix(
    unlist(
      mean_FWB_by_fpl.list), nrow=length(mean_FWB_by_fpl.list), byrow=TRUE)) 

colnames(mean_FWB_by_fpl.list2) <- c("TK")

mean_FWB_by_fpl.list2$meanfwB <- c("200%+ FPL", "100% - 199% FPL", ">100% FPL", "Overall Mean")

mean_FWB_by_fpl2 <- select(mean_FWB_by_fpl.list2, meanfwB:TK)


#Plots

fplLevelsFWB <- ggplot(mean_FWB_by_fpl2) +
  geom_col(aes(
    x=factor(
      meanfwB, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=TK, group = 1, color = meanfwB, fill = meanfwB))
print(fplLevelsFWB)

fplLevelsFWB <- fplLevelsFWB +
  theme(legend.position = "bottom") +
  labs(title = "Financial Well-Being Scale Score", x = "", y = "")
print(fplLevelsFWB)

# Mean FS Score measures by Federal Poverty Level

mean_FSscore <- mean(as.integer(df3$FSscore))
print(mean_FSscore)

df3_mean_FSscore.processed <- 
  df3 %>%
  select(FSscore)
df3_meanFSscore.processed <-
  df3 %>%
  summarise(
    mean_FSscore = mean(as.integer(df3$FSscore)))

meanFSscore_byFPL1 <- 
  df3 %>%
  filter(fpl == "200%+ FPL") %>%
  select(FSscore)
meanFSscore_byFPL1 <- 
  meanFSscore_byFPL1 %>%
  summarise(
    mean_FSscore = mean(as.integer(meanFSscore_byFPL1$FSscore)))

meanFSscore_byFPL2 <- 
  df3 %>%
  filter(fpl == "100%-199% FPL") %>%
  select(FSscore)
meanFSscore_byFPL2 <- 
  meanFSscore_byFPL2 %>%
  summarise(
    mean_FSscore = mean(as.integer(meanFSscore_byFPL2$FSscore)))

meanFSscore_byFPL3 <- 
  df3 %>%
  filter(fpl == "<100% FPL") %>%
  select(FSscore)
meanFSscore_byFPL3 <- 
  meanFSscore_byFPL3 %>%
  summarise(
    mean_FSscore = mean(as.integer(meanFSscore_byFPL3$FSscore)))

mean_FS_by_fpl.list <-
  list(
    meanFSscore_byFPL3, meanFSscore_byFPL2, meanFSscore_byFPL1, df3_meanFSscore.processed)

Reduce(function(x, y) merge(x, y, all=TRUE), mean_FS_by_fpl.list)

mean_FS_by_fpl.list2 <- data.frame(
  matrix(
    unlist(
      mean_FS_by_fpl.list), nrow=length(mean_FS_by_fpl.list), byrow=TRUE)) 

colnames(mean_FS_by_fpl.list2) <- c("TK")

mean_FS_by_fpl.list2$meanfS <- c("200%+ FPL", "100% - 199% FPL", ">100% FPL", "Overall Mean")

mean_FS_by_fpl2 <- select(mean_FS_by_fpl.list2, meanfS:TK)


#Plots

fplLevelsFS <- ggplot(mean_FS_by_fpl2) +
  geom_col(aes(
    x=factor(
      meanfS, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=TK, group = 1, color = meanfS, fill = meanfS))
print(fplLevelsFS)

fplLevelsFS <- fplLevelsFS +
  theme(legend.position = "bottom") +
  labs(title = "Financial Skill Scale Score", x = "", y = "")
print(fplLevelsFS)


# Mean LM Score measures by Federal Poverty Level

mean_LMscore <- mean(as.integer(df3$LMscore))
print(mean_LMscore)

df3_mean_LMscore.processed <- 
  df3 %>%
  select(LMscore)
df3_meanLMscore.processed <-
  df3 %>%
  summarise(
    mean_LMscore = mean(as.integer(df3$LMscore)))

meanLMscore_byFPL1 <- 
  df3 %>%
  filter(fpl == "200%+ FPL") %>%
  select(LMscore)
meanLMscore_byFPL1 <- 
  meanLMscore_byFPL1 %>%
  summarise(
    mean_LMscore = mean(as.integer(meanLMscore_byFPL1$LMscore)))

meanLMscore_byFPL2 <- 
  df3 %>%
  filter(fpl == "100%-199% FPL") %>%
  select(LMscore)
meanLMscore_byFPL2 <- 
  meanLMscore_byFPL2 %>%
  summarise(
    mean_LMscore = mean(as.integer(meanLMscore_byFPL2$LMscore)))

meanLMscore_byFPL3 <- 
  df3 %>%
  filter(fpl == "<100% FPL") %>%
  select(LMscore)
meanLMscore_byFPL3 <- 
  meanLMscore_byFPL3 %>%
  summarise(
    mean_LMscore = mean(as.integer(meanLMscore_byFPL3$LMscore)))

mean_LM_by_fpl.list <-
  list(
    meanLMscore_byFPL3, meanLMscore_byFPL2, meanLMscore_byFPL1, df3_meanLMscore.processed)

Reduce(function(x, y) merge(x, y, all=TRUE), mean_LM_by_fpl.list)

mean_LM_by_fpl.list2 <- data.frame(
  matrix(
    unlist(
      mean_LM_by_fpl.list), nrow=length(mean_LM_by_fpl.list), byrow=TRUE)) 

colnames(mean_LM_by_fpl.list2) <- c("TK")

mean_LM_by_fpl.list2$meanLM <- c("200%+ FPL", "100% - 199% FPL", ">100% FPL", "Overall Mean")

mean_LM_by_fpl2 <- select(mean_LM_by_fpl.list2, meanLM:TK)


#Plots

fplLevelsLM <- ggplot(mean_LM_by_fpl2) +
  geom_col(aes(
    x=factor(
      meanLM, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=TK, group = 1, color = meanLM, fill = meanLM))
print(fplLevelsLM)

fplLevelsLM <- fplLevelsLM +
  theme(legend.position = "bottom") +
  labs(title = "Lusardi & Mitchell Financial Knowledge Scale Score", x = "", y = "")
print(fplLevelsLM)

# Mean KH Score measures by Federal Poverty Level

mean_KHscore <- mean(as.integer(df3$KHscore))
print(mean_KHscore)

df3_mean_KHscore.processed <- 
  df3 %>%
  select(KHscore)
df3_meanKHscore.processed <-
  df3 %>%
  summarise(
    mean_KHscore = mean(as.integer(df3$KHscore)))

meanKHscore_byFPL1 <- 
  df3 %>%
  filter(fpl == "200%+ FPL") %>%
  select(KHscore)
meanKHscore_byFPL1 <- 
  meanKHscore_byFPL1 %>%
  summarise(
    mean_KHscore = mean(as.integer(meanKHscore_byFPL1$KHscore)))

meanKHscore_byFPL2 <- 
  df3 %>%
  filter(fpl == "100%-199% FPL") %>%
  select(KHscore)
meanKHscore_byFPL2 <- 
  meanKHscore_byFPL2 %>%
  summarise(
    mean_KHscore = mean(as.integer(meanKHscore_byFPL2$KHscore)))

meanKHscore_byFPL3 <- 
  df3 %>%
  filter(fpl == "<100% FPL") %>%
  select(KHscore)
meanKHscore_byFPL3 <- 
  meanKHscore_byFPL3 %>%
  summarise(
    mean_KHscore = mean(as.integer(meanKHscore_byFPL3$KHscore)))

mean_KH_by_fpl.list <-
  list(
    meanKHscore_byFPL3, meanKHscore_byFPL2, meanKHscore_byFPL1, df3_meanKHscore.processed)

Reduce(function(x, y) merge(x, y, all=TRUE), mean_KH_by_fpl.list)

mean_KH_by_fpl.list2 <- data.frame(
  matrix(
    unlist(
      mean_KH_by_fpl.list), nrow=length(mean_KH_by_fpl.list), byrow=TRUE)) 

colnames(mean_KH_by_fpl.list2) <- c("TK")

mean_KH_by_fpl.list2$meanKH <- c("200%+ FPL", "100% - 199% FPL", ">100% FPL", "Overall Mean")

mean_KH_by_fpl2 <- select(mean_KH_by_fpl.list2, meanKH:TK)


#Plots

fplLevelsKH <- ggplot(mean_KH_by_fpl2) +
  geom_col(aes(
    x=factor(
      meanKH, level=c("Overall Mean", ">100% FPL", "100% - 199% FPL", "200%+ FPL")), y=TK, group = 1, color = meanKH, fill = meanKH))
print(fplLevelsKH)

fplLevelsKH <- fplLevelsKH +
  theme(legend.position = "bottom") +
  labs(title = "Knoll & Houts Financial Knowledge Scale Score", x = "", y = "")
print(fplLevelsKH)


```


# Analysis

## Question 1: What observations can be inferred by viewing each of the finanicial well-being, skills and knowledge measures that have been charted visually?

# Figure 1 displays the average numerical response level by survey participants for each of the Subjective Well-Being (SWB) Questions, as three separate graphs (1 graph per survey question). Participants were asked to report their sentiments, using a scale ranging from 1 to 7 (with "1" meaning "Strongly disagree" and "7" meaning "Strongly Agree") after hearing each of the following statements: (1) "I am satisfied with my life"; (2) "I am optimistic about my future"; (3) "If I work hard today, I will be more successful in the future."

# For Question 1 -- "I am satisfied with my life" -- the average self-reported SWB assessment was approximately 5.79, meaning that as a sample, participants were likely on average to report firm agreement with this statement. For Question 2 -- "I am optimistic about my future" -- the average self-reported SWB assessment was approximately 6.12, meaning that as a sample, participants were likely on average to report firmly agree with this statement. For Question 3 -- "If I work hard today, I will be more successful in the future" -- participants reported the highest level of agreement, at an average SWB level of 6.34.

# Comparing each respective overall average to the averages that were calculated by  disaggregating partcipant responses by Federal Poverty Level (FPL) status. Interestingly, for each of the three questions, individuals with earnings either at the FPL, or up to 199% above it, self-reported the highest SWB levels.

# Figure 2 displays both the overall and by-FPL measures of Financial Well-Being Scale Score (FWB). Recorded scores ranged from scores of 14 to 95; the full score scale ranged from 0 to 100. Approximately 86 different numerical scores were recorded. These scores were accumulated from responses to ten different questions and participant responses related to financial well-being. For Figure 2, the overall average is approximately 43.01. The average FWB for particpants earning less than 100% of the FPL measured the highest, at 45.65; at 100% to 199% of the FPL, and 200% and above the FPL, FWB measures were approximately 36.23 and 32.31, respectively.

# Similar trends in Figure 2 can be seen in Figure 3, which displays the average Financial Skill (FS) Scale Score overall and by FPL. Measured on a similar 0 to 100 scale as the FWB, scores are generally expected to fall within a middle threshold; extremely high or extremely low number scores are uncommon. The overall average scale score was 42.30; 43.45 for surveyed individuals whose earnings were lower than the FPL; 38.81 for earners between 100% and 199% of the FPL; and 38.33 for participants earning 200% above the FPL or higher.

# Figure 4 displays average Lusardi and Mitchell (LM) Financial Knowledge Scale scores by overall averages as well as averages by FPL. Scores are calculated from a financial literacy questionnaire, scored on a scale ranging from 0 to 9, that assesses participants' knowledge of compound interest, risk diversification, and inflation concepts based on questions that are answered correctly or incorrectly. Once again, similar trends emerged. The overall LM score average score was 3.51, with a slight uptick in average score for individuals who earned at lower than the FPL (3.63), a decrease for people who earned between 100% and 199% of the FPL (3.26), and a further decrease for individuals who earned above 200% and higher than the FPL (2.91).

# Lastly, Figure 5 displays average Knoll and Houts (KH) Financial Knowledge Scale scores, which has the similar purpose and structure as Lusardi's and Mitchell's scale but add questions about interest and the time value of money. Yet again, similar trends occured. The overall KH score average score was 6.11, with an uptick in average score for individuals who earned at lower than the FPL (6.62), a sizable decrease for people who earned between 100% and 199% of the FPL (4.80), and a further decrease for individuals who earned above 200% and higher than the FPL (4.10).

\newpage

# Summary and Conclusions

# The Consumer Financial Protection Bureau's National Financial Well-Being Survey to examine relationships between self-reported measures of financial well-being to three different assessments of financial literacy, with an emphasis on potential differences that occurred for earnings below, at and above the Federal Poverty Level. Self-reported measures of financial well-being were high among survey participants. On average, respondents are satisifed with their lives, are optimistic about their futures, and believe that hard work will lead to good future financial outcomes. Participants' Financial Well-Being scores fall within average ranges, both overall and across different earnings tiers by Federal Poverty Level. Participants who earned less than the Federal Poverty Level threshold scored highest on this scale, followed by individuals earning between 100% and 199% of the Federal Poverty Level and respondents who earned twice as much as the Federal Poverty level or above, respectively. Similar rankings of peformance occured for each of the financial literacy scores that were assessed.

# Overall, from the data analysis at hand it appears that measurements of financial well-being and financial literacy differed little between teh different income groups used by the CFPB and for this dataset. This research, howver, is not exhaustive; in fact, it is preliminary. With further time to investigate this CFPB data more thoroughly, a number of statistical calculations would have been possible. For example, further visualizations could have examined if that were any statistically significant correlations between the measures studied for this experiment, including statistical analysis such as t-tests, ANOVA and the like. Further graphical disaggregations may have also been possible. For example, further analysis may have allowed the researchers for this experiment to explore differences in expressions of financial literacy that could be broken down not only by Federal Poverty Level, but also by the numerical value of participants' answers to each of the Subjective Well-Being questions. Meaning, it may be useful to assess the financial literacy of a person or group to earned less than the Federal Poverty threshold, held high financial knowledge according to the scales used for this survey, yet may have reported to subjectively feel that working hard may not bring future financial success. Or, some other combination of earnings, subjective financial outlooks and financial literacy might possibiliy emerge outside of the trends interpreted so far. 

\newpage

# References
<add references here if relevant, otherwise delete this section> 
